---
title: "Visualize algea bloom as wave"
author: "Miguel D. Mahecha, Guido Kraemer with initial suggestions from Veronika Grupp"
output:
  html_document:
    self_contained: yes
---


# Summary

A documented analysis of the 2022 algea bloom in the river Oder based on remote
sensing estimates of chlorophyll concentrations.

The idea of the script is the following:

* Load the river network as offered by INSPIRE
* Transform the river of interest into a spatial network using sfnetwork
* Match the following data to the river network:
  * The distance for every point to the mouth of the river
  * The S2 estimated chlorophyll concentrations and their coordinates
  * the historical reanalysis discharge data from EFAS
* Additionally we add discharge observation data for the stations
  Hohensaaten-Finow and Eisenhüttenstadt.
* Visualize the data

# Analysis

```{r}
#| message: FALSE
#| warning: FALSE
library(tidyverse)
library(sf)
library(sfnetworks)
library(tidygraph)
library(igraph)
library(lwgeom)
library(viridis)
library(lubridate)
```

# River network and distances to the mouth
## Read the river network data and select rivers of interest

Read in the INSPIRE data and extract all the information we need.

```{r}
inspire_data_path <- "data/RWB_2016_ManagementRestrictionOrRegulationZone_2020_L.gml+xml"
river_net_full <- st_read(inspire_data_path)
river_net_full <- select(river_net_full, gml_id, text, geometry)
```

The following rivers are potentially of interest (Oder and tributaries) We give
the names in Polish exactly as in the INSPIRE file and in parenthesis the
English and German names

* ODRA (Oder, Oder)
* NYSA KŁODZKA (Eastern Neisse, Glatzer Neiße)
* NYSA ŁUŻYCKA (Lusatian Neiss,  Lausitzer Neiße)
* WARTA (Warta, Warthe)
* KANAŁ GLIWICKI Z KŁODNICĄ OD KOZŁÓWKI DO DRAMY
* KANAŁ GLIWICKI (Gleiwitzer Kanal)


```{r}
river_filter_terms <- c(
  "ODRA",
  "NYSA KŁODZKA",
  "NYSA ŁUŻYCKA",
  "NYSA ŁUZYCKA",
  "WARTA",
  "ZB. PORAJ",
  "KANAŁ GLIWICKI",
  "KŁODNICA"
)

river_filter_idx <- lapply(river_filter_terms, grep, x = river_net_full$text)
river_filter_idx <- do.call(c, river_filter_idx)

river_net <- river_net_full[river_filter_idx, ]

#### This is the missing piece for the WARTA, we add it manually and get the
#### number of components from 2 to 1. Coordinates are extracted manually using
#### qgis
warta_missing_piece <- matrix(c(19.37629795, 50.59313292,
                                19.37595002, 50.59261401), 2, 2, byrow = TRUE) %>%
  list %>%
  st_multilinestring %>%
  st_sfc(crs = st_crs(river_net)) %>%
  st_sf

river_net <- bind_rows(river_net, warta_missing_piece)

ggplot(river_net) + geom_sf() + theme_bw()
```

## Transform rivers of interest to network

In order to calculate the distance of each point to the mouth of the river we
need to transform our river network into a connected graph.

The sfnetwork class allows us the retain the geographical information of the
graph.

```{r}

# The data we read in come as MULTILINESTRING
river_net_linestring <- st_cast(river_net, "LINESTRING")
river_net_point <- st_cast(river_net_linestring, "POINT")
# as_sfnetwork only uses enpoints as nodes, here we cut our data up so that each
# linestring consists of only two points and therefore every point becomes an
# endpoint
river_net_collections <- lwgeom::st_split(river_net_linestring, river_net_point)

# here each geometry is a GEOMETRYCOLLECTION of many LINESTRING, we need to
# separate them.

river_net_rows <- list()
for (i in 1:nrow(river_net_collections)) {
  river_net_collection <- river_net_collections[i, ]
  ### using c(...) makes the whole thing flat and not nested
  river_net_rows <- c(
    river_net_rows,
    lapply(river_net_collection$geometry[[1]], function(x) {
      river_net_collection$geometry <- st_sfc(x)
      river_net_collection
    }))
}


### a list of length 51000
length(river_net_rows)
### takes ~20s
## rows <- bind_rows(rows)
### this is a much faster version and should give the same result
river_net_rows_gml_id <- vapply(river_net_rows, function(x) x$gml_id, "")
river_net_rows_text <- vapply(river_net_rows, function(x) x$text, "")
## sapply doesn't work here...
river_net_rows_geometry <- lapply(river_net_rows, function(x) x$geometry)
river_net_rows_geometry <- do.call(c, river_net_rows_geometry)
st_crs(river_net_rows_geometry) <- st_crs(river_net)
river_net_rows <- st_sf(gml_id = river_net_rows_gml_id,
                        text = river_net_rows_text,
                        geometry = river_net_rows_geometry,
                        crs = st_crs(river_net))

### a sf object with 51000 LINESTRING
river_net_rows

### Now we create the actual network and add the length of each segment as
### weights for a later distance calculation.
river_sfnet = as_sfnetwork(river_net_rows, directed = FALSE) %>%
  activate("edges") %>%
  mutate(weight = edge_length())

### Sanity check, this must be a single component!
stopifnot(components(river_sfnet)$no == 1)
stopifnot(is_connected(river_sfnet))

### We do not make any plots here because plotting becomes really slow due to
### the many tiny LINESTRING objects
```

## Distance calculations

We add the distance to the river mouth as a property to each vertex of the graph

```{r}
### choose the northern most point as the mouth of the river
idx_river_mouth <- river_sfnet %>%
  activate("nodes") %>%
  st_coordinates %>% {
    .[, 2]
  } %>%
  which.max

### distances from the mouth of the river to all points
distance_from_mouth <- igraph::distances(river_sfnet, idx_river_mouth) %>%
  as.vector
V(river_sfnet)$distance_from_mouth <- distance_from_mouth

### Sanity plot
river_sfnet %>%
  activate("nodes") %>% {
    coords <- st_coordinates(.)
    tibble(x = coords[, "X"], y = coords[, "Y"],
           distance_from_mouth = distance_from_mouth)
  } %>%
  ggplot() +
  aes(x = x, y = y, color = distance_from_mouth) +
  geom_point()
```

# Match chlorophyll observations to the river network
## Read chlorophyll concentrations

These data have been generated by Kerstin from Brockmann Consult using the
coordinates from the same INSPIRE data. The data are organized conceptually as
tidy data with one observation by location and time.

```{r}
# Data from Brockmann Consult
chlorophyll_data_path <- "data/chlorophyll_data.csv"

# Note that we did a manual intervention removing white spaces from river station names

# read into tibble
# note that we defined column types manually as read_table makes wrong guesses
chlorophyll_data <- read_csv(
  chlorophyll_data_path,
  col_types = list(
    Name = col_character(),
    Date = col_date(format = ""),
    Longitude = col_double(),
    Latitude = col_double(),
    Chlorophyll_mean = col_double(),
    Chlorophyll_sigma = col_double(),
    num_passes = col_double()
  ),
  comment = "#"
)

# we only care about locations where we do have valid chl observations and skip the rest
chlorophyll_data <- chlorophyll_data %>%
  drop_na(Chlorophyll_mean) %>%
  rename(X = Longitude, Y = Latitude)

chlorophyll_data_sf <- st_as_sf(chlorophyll_data, coords = c("X", "Y"))
st_crs(chlorophyll_data_sf) <- st_crs(river_net)

ggplot() +
  geom_sf(data = river_net) +
  geom_point(data = chlorophyll_data,
             aes(x = X, y = Y, color = Chlorophyll_mean))

```


## Match distance to mouth to our chlorophyll data

```{r}
### Get the index of the nearest point in the river network to the chlorophyll
### data
nearest_vertex_idx <- river_sfnet %>%
  V %>%
  `$`(geometry) %>%
  st_nearest_feature(chlorophyll_data_sf, .)

### We take the distance to the river mouth from the river network an put it
### into the chlorophyll data
chlorophyll_data_sf$distance_from_mouth <-
  V(river_sfnet)$distance_from_mouth[nearest_vertex_idx]
```

# Add cities and towns

We add some cities to make the maps easier to read. Data taken from Open Street Map.

```{r}

oder_cities <- st_read("data/oder_cities.geojson") %>%
  st_transform(st_crs(river_sfnet)) %>%
  select(name)
oder_towns <- st_read("data/oder_towns.geojson") %>%
  st_transform(st_crs(river_sfnet)) %>%
  select(name)

oder_cities <- oder_cities %>%
  filter(name != "Ostrava")

oder_towns <- oder_towns %>%
  filter(name == "Frankfurt (Oder)")

oder_cities <-
  rbind(oder_cities, oder_towns)


nearest_vertex_idx <- river_sfnet %>%
  V %>%
  `$`(geometry) %>%
  st_nearest_feature(oder_cities, .)

oder_cities$distance_from_mouth <-
  V(river_sfnet)$distance_from_mouth[nearest_vertex_idx]

nearest_vertex_idx <- river_sfnet %>%
  V %>%
  `$`(geometry) %>%
  st_nearest_feature(oder_towns, .)

oder_towns$distance_from_mouth <-
  V(river_sfnet)$distance_from_mouth[nearest_vertex_idx]

oder_towns$distance_from_mouth

ggplot() +
  geom_sf(data = river_net) +
  geom_sf(data = oder_cities, aes(color = distance_from_mouth))
  ## geom_sf(data = oder_towns, aes(color = distance_from_mouth))
```

# Chlorophyll plots
## Chlorophyll vs. Distance to mouth

Plot the chlorophyll concentration in function to the mouth of the Oder by time
slice.

```{r}
library(ggspatial)

datecuts <- c("2022-07-01", "2022-07-21", "2022-08-01",
              "2022-08-05", "2022-08-09", "2022-08-16",
              "2022-08-19") %>%
  as.Date

startdate <- chlorophyll_data_sf$Date %>% min
enddate <- chlorophyll_data_sf$Date %>% max


fmt <- function(x) format(x, "%d/%m")
datecutlabels <- paste(fmt(datecuts[-length(datecuts)]),
                       fmt(datecuts[-1] - as.difftime("24:00:00")),
                       sep = "–")

chlorophyll_data_sf %>%
  mutate(date_cut = cut(Date, datecuts, labels = datecutlabels)) %>%
  ggplot() +
  aes(x = distance_from_mouth / 1000, y = Chlorophyll_mean) +
  geom_point(color = "#31a354") +
  geom_smooth(span = 0.2, color = "#fc8d62", fill = NA) +
  geom_vline(xintercept = oder_cities$distance_from_mouth / 1000) +
  geom_text(data = oder_cities, aes(x = distance_from_mouth / 1000, y = Inf, label = name),
            vjust = 1.1, hjust = 0, angle = -90) +
  ylim(0, 400) +
  scale_x_reverse() +
  labs(y = expression("Chlorophyll concentration [" * mu * "g/l]"),
       x = "Distance to river mouth [km]") +
  facet_wrap(vars(date_cut), nrow = 4) +
  theme_minimal()
```

## Hexbin plot ##

Just plotting chlorophyll concentrations for each pixel will result in
overlapping pixels, therefore we need to bin the data before plotting.

```{r, fig.height = 3}

cddf <- chlorophyll_data %>%
  mutate(date_cut = cut(Date, datecuts, labels = datecutlabels))

ocdf <- cbind(oder_cities, date_cut = factor(levels(cddf$date_cut)[6],
                                             levels = levels(cddf$date_cut)))

scale_data <- data.frame(date_cut = cddf$date_cut[1],
                         location = "bl")

cddf %>%
  ggplot() +
  geom_sf(data = river_net, color = "gray90") +
  stat_summary_hex(aes(x = X, y = Y, z = Chlorophyll_mean),
                   fun = mean, na.rm = TRUE) +
  geom_sf(data = oder_cities) +
  geom_segment(data = ocdf,
               aes(xend = Inf, yend = after_stat(y), geometry = geometry),
               color = "gray70", stat = "sf_coordinates") +
  geom_label(data = ocdf,
             aes(label = name, geometry = geometry),
             x = Inf, hjust = 1, label.size = NA,
             color = "gray70", stat = "sf_coordinates") +
  scale_fill_distiller(expression("Chl. [" * mu * "g/l]"),
                       palette = "YlGn",
                       direction = 1) +
  facet_wrap(vars(date_cut), nrow = 1) +
  ggspatial::annotation_scale(data = scale_data,
                              aes(location = location),
                              text_col = "gray70",
                              line_col = "gray80",
                              bar_cols = c("gray80", "white")) +
  theme_void() +
  theme(panel.spacing.x = unit(-3, "cm"),
        strip.text = element_text(hjust = 0))

```

# Discharge data EFAS
## Data ##
Discharge comes from the EFAS 6 hourly data. We are using the historical Version
4.0 data of river discharge in the last 6h data.

- The API call can be found in `download_discharge_data.py`
- The dataset can be found here: https://cds.climate.copernicus.eu/cdsapp#!/dataset/efas-historical

## functions for accessing EFAS data

```{r}
library(ncdf4)

discharge_crs <- "+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs"

get_latslons <- function(x, reproj = FALSE) {
  # get latitude longitude mask through manual reprojection because the
  # latitude/longitude data is not well defined. Testing it there are some clear
  # discrepancies. They are probably using a different ellipsoid.
  if (reproj) {
    xx <- ncvar_get(x, "x")
    yy <- ncvar_get(x, "y")
    xxyy_grid <- expand.grid(x = xx, y = yy)
    grid_lambert <- st_as_sf(xxyy_grid, coords = c("x", "y"),
                             crs = st_crs(discharge_crs))
    res <- st_transform(grid_lambert, st_crs(river_net))
    # alternatively just take the latitudes and longitudes as they are provided
    # by the data
  } else {
    xx <- ncvar_get(x, "longitude") %>% as.vector
    yy <- ncvar_get(x, "latitude") %>% as.vector
    xxyy_grid <- data.frame(x = xx, y = yy)
    res <- st_as_sf(xxyy_grid,
                    coords = c("x", "y"),
                    crs = st_crs(river_net))
  }
  # HACK: In the end neither of the above methods really worked so we just set a
  # threshold and remove everything below it. In order to keep indexing still
  # working we just move the points to a far away location.

  # the data does not really match the river data, so here we remove everything
  # with < 8 m3s-1 on the first time step.
  d <- ncvar_get(x, "dis06", count = c(-1, -1, 1))
  dmask <- as.vector(d > 8)
  res$geometry[!dmask] <- st_point(c(0, 0))
  return(res)
}

get_time <- function(x) {
  # [time]
  t_axis <- ncvar_get(x, "time")
  as.POSIXct(t_axis, origin = "1970-01-01")
}

get_data <- function(x) {
  # [x, y, time]
  ncvar_get(x, "dis06")
}

mask_margin <- function(x, mask) {
  # x: array [x, y, time]
  # mask: matrix [obs, linear[x, y]]
  dx <- dim(x)
  dim(x) <- c(dx[1] * dx[2], dx[3])
  x[mask, ]
}
```

## Create mask for discharge data ##

The discharge data is relatively large when transformed into a data.frame in
long form (i.e. with columns [location, time, discharge]). We need to create a
mask first to filter the spatial locations we actually need.

```{R}
discharge_file_names <- dir("data", "discharge_.*_cropped\\.nc$", full.names = TRUE)
discharge_nc <- lapply(discharge_file_names, nc_open)

discharge_nx <- ncvar_get(discharge_nc[[1]], "x") %>% length
discharge_ny <- ncvar_get(discharge_nc[[1]], "y") %>% length
discharge_nyear <- length(discharge_nc)

pixel_locations <- get_latslons(discharge_nc[[1]])
st_crs(pixel_locations) <- st_crs(chlorophyll_data_sf)

discharge_location_idxs_lin <- st_nearest_feature(chlorophyll_data_sf, pixel_locations)
discharge_location_idxs_arr <- arrayInd(discharge_location_idxs_lin, c(discharge_nx, discharge_ny))

discharge_nspace <- length(discharge_location_idxs_lin)
```

## Load discharge data ##
### Raw data ###

Connecting the data directly with the chlorophyll data would be overkill because
we would blow up the number of rows by a factor equal to the number of time
steps. There are many duplicated indices in `discharge_location_idxs_lin` for
the calculations we only need unique locations in space.

```{r}

discharge_location_idxs_lin_uniq <-
  sort(unique(discharge_location_idxs_lin))

discharge_data <- discharge_nc %>%
  lapply(function(x) {
    x2 <- get_data(x)
    mask_margin(x2, discharge_location_idxs_lin_uniq)
  })
discharge_space <- pixel_locations$geometry[discharge_location_idxs_lin_uniq]
# The time vectors can have different lengths! maybe a leap year thing?
discharge_time <- lapply(discharge_nc, get_time)

```

put everything into a data frame

```{r}

# it would be much better to aggregate *before* transforming this into a
# data.frame. Performance is not great but okay.
discharge_df_list <- lapply(seq_len(discharge_nyear), function(i) {
  # [space, time]
  d <- discharge_data[[i]]
  # [time]
  t <- rep(discharge_time[[i]], each = dim(d)[1])
  # [space]
  s <- rep(discharge_space, dim(d)[2])
  # [space]
  i <- rep(discharge_location_idxs_lin_uniq, dim(d)[2])

  tibble(discharge = as.vector(d), time = t, geometry = s, location_idx = i)
})

discharge_df <- bind_rows(discharge_df_list)
```

aggregate and perform MSC and anomaly calculations.

```{r}

# aggregate by day
discharge_df <- discharge_df %>%
  # as.Date(xxxxx 00:00h) seems to subtract one day
  group_by(Date = as.Date(floor_date(time, "days")) + 1, geometry, location_idx) %>%
  dplyr::summarize(discharge = mean(discharge)) %>%
  ungroup

# mean seasonal cycle
discharge_df <- discharge_df %>%
  group_by(yday = yday(Date), location_idx) %>%
  mutate(discharge_msc = mean(discharge)) %>%
  ungroup %>%
  mutate(discharge_anomaly = discharge - discharge_msc)

# cumulative discharge and discharge anomaly since with zero at june 1
discharge_df <- discharge_df %>%
  group_by(year = year(Date), location_idx) %>%
  arrange(Date) %>%
  mutate(
    discharge_anom_cum_june =
      cumsum(discharge_anomaly) %>%
      { . - .[month(Date) == 6 & mday(Date) == 1] },
    discharge_cum_june =
      cumsum(discharge) %>%
      { . - .[month(Date) == 6 & mday(Date) == 1] }
  ) %>%
  ungroup

# percentile of cumulative discharge
discharge_df <- discharge_df %>%
  group_by(yday = yday(Date)) %>%
  mutate(
    discharge_anom_cum_june_perc =
      rank(discharge_anom_cum_june) %>%
      { . / max(.) * 100 },
    discharge_cum_june_perc =
      rank(discharge_cum_june) %>%
      { . / max(.) * 100 },
    discharge_perc =
      rank(discharge) %>%
      { . / max(.) * 100}
  ) %>%
  ungroup
```

### Discharge Plots

do some plots for sanity check.

```{r}

ggplot(discharge_df) +
  geom_line(aes(x = Date, y = discharge, group = location_idx, color = "discharge"), alpha = 0.1) +
  geom_line(aes(x = Date, y = discharge_msc, group = location_idx, color = "msc"), alpha = 0.1)

ggplot(discharge_df) +
  aes(x = Date, y = discharge_anomaly, group = location_idx) +
  geom_line(alpha = 0.5)

hist(discharge_df$discharge, main = "histogram of discharge")
hist(discharge_df$discharge_msc, main = "histogram of discharge msc")

```

## Discharge at specific locations ##

In the end we are only interested in the discharge at some specific locations

River discharge a selected cities along the Oder

```{r}
discharge_city_idxs <- discharge_location_idxs_lin_uniq[(st_nearest_feature(oder_cities, discharge_space))]

discharge_df %>%
  filter(location_idx %in% discharge_city_idxs) %>%
  left_join(tibble(city = oder_cities$name, location_idx = discharge_city_idxs)) %>%
  ggplot() +
  aes(x = Date, y = discharge_anomaly) +
  geom_line() +
  facet_wrap(vars(city), scales = "free_y")
```

### Zoom into 2022: ###

We are mostly interested in the algae bloom in summer 2022

```{r}
discharge_df %>%
  filter(location_idx %in% discharge_city_idxs) %>%
  filter(year(Date) == 2022) %>%
  left_join(tibble(city = oder_cities$name, location_idx = discharge_city_idxs)) %>%
  ggplot() +
  aes(x = Date, y = discharge_anomaly) +
  geom_line() +
  facet_wrap(vars(city), scales = "free_y")

discharge_df %>%
  filter(location_idx %in% discharge_city_idxs) %>%
  filter(year(Date) == 2022) %>%
  filter(month(Date) %in% 5:8) %>%
  left_join(tibble(city = oder_cities$name, location_idx = discharge_city_idxs)) %>%
  ggplot() +
  geom_line(aes(x = Date, y = discharge, color = "discharge")) +
  geom_line(aes(x = Date, y = discharge_msc, color = "msc")) +
  facet_wrap(vars(city), scales = "free_y")
```

## Cumulative discharge and discharge anomaly ##

prepare the cumulative discharge data

```{r}

discharge_df_cities <- discharge_df %>%
  filter(location_idx %in% discharge_city_idxs) %>%
  filter((month(Date) %in% 6:7) | (month(Date) == 8 & mday(Date) < 16)) %>%
  left_join(tibble(city = oder_cities$name,
                   location_idx = discharge_city_idxs)) %>%
  # "normalize" Date by year, so that we can compare different years
  mutate(Date0 = as.Date(yday(Date), origin = "2021-12-31"))
```

read in the measured data
```{r}
library(zoo)
data_eh_path <- "data/6030000_Eisenhuettenstadt_Q_TagWerte.csv"
data_hf_path <- "data/6030800_Hohensaaten-Finow_Q_TagWerte.csv"

read_measured_data <- function(path) {
  l <- readLines(path, n = 4)
  city <- l[2] %>% sub("#station_name;", "", ., fixed = TRUE)
  lat <- l[3] %>% sub("#station_latitude;", "", ., fixed = TRUE)
  lon <- l[4] %>% sub("#station_longitude;", "", ., fixed = TRUE)
  df <- read_delim(
    path,
    col_names = c("time", "discharge"),
    comment = "#",
    delim = ";",
    locale = readr::locale(tz = "Europe/Berlin"),
    show_col_type = FALSE
  )
  df <- df %>%
    mutate(X = lon,
           Y = lat,
           city = city,
           Date = lubridate::as_date(time),
           Date0 = as.Date(yday(Date), origin = "2021-12-31"),
           yday = yday(Date),
           year = year(Date)) %>%
    ## year 1962 starts in september
    filter(year != 1962) %>%
    ## ## filter to match the simulated date
    ## filter(year >= 1991) %>%
    ## we also need to do some gapfilling
    mutate(discharge = na.approx(discharge, na.rm = FALSE))

  ## na.approx doesn't fill edges!
  if(is.na(df$discharge[length(df$discharge)])) {
    df$discharge[length(df$discharge)] <- df$discharge[length(df$discharge) - 1]
  }
  if(is.na(df$discharge[1])) {
    df$discharge[2]
  }

  stopifnot(!anyNA(df$discharge))

  ## calculate discharge msc and anomaly
  df <- df %>%
    group_by(yday) %>%
    mutate(discharge_msc = mean(discharge, na.rm = TRUE)) %>%
    ungroup() %>%
    mutate(discharge_anomaly = discharge - discharge_msc)


  ## calculate cumulative discharge and anomaly
  df <- df %>%
    group_by(year) %>%
    arrange(Date) %>%
    ## we just calculate the cumulative sum over the entire year and
    ## substract the date where we want to start (1/6)
    mutate(
      discharge_anom_cum_june = cumsum(discharge_anomaly),
      discharge_anom_cum_june = discharge_anom_cum_june - discharge_anom_cum_june[month(Date) == 6 & mday(Date) == 1],
      discharge_cum_june = cumsum(discharge),
      discharge_cum_june = discharge_cum_june - discharge_cum_june[month(Date) == 6 & mday(Date) == 1]
    ) %>%
    ungroup()
  ## filter out the date range and transform int sf object
  df <- df %>%
    filter((month(Date) %in% 6:7) | (month(Date) == 8 & mday(Date) < 16)) %>%
    st_as_sf(coords = c("X", "Y"), crs = st_crs(river_net))
  return(df)
}

discharge_df_cities2 <-
  bind_rows(read_measured_data(data_eh_path),
            read_measured_data(data_hf_path),
            discharge_df_cities)

discharge_df_cities2 %>%
  filter(yday(Date) == yday(max(Date))) %>%
  arrange(-year(Date))

```

plot cumulative discharge data

```{r, fig.height = 12, fig.width = 12}
library(ggrepel)

ggplot() +
  aes(x = Date0,
      y = discharge_cum_june,
      color = year == 2022,
      group = year,
      label = year,
      linewidth = as.numeric(year == 2022)) +
  geom_line(data = discharge_df_cities2) +
  scale_color_discrete(type = c("grey50", "red")) +
  geom_text_repel(
    data = discharge_df_cities2 %>% filter(yday(Date) == yday(max(Date))),
    segment.colour = "gray80",
    size = 3,
    direction = "y",
    nudge_x = 20,
    hjust = 0
  ) +
  scale_linewidth(range = c(0.5, 1)) +
  facet_wrap(vars(city), scales = "free_y") +
  labs(y = "Cumulative discharge [m3s-1]",
       x = "Date",
       color = "Year",
       fill = "Year") +
  theme_minimal() +
  theme(legend.justification = c(1, 0),
        legend.position = "none")

ggplot() +
  aes(x = Date0, y = discharge_anom_cum_june,
      color = year(Date) == 2022,
      group = year(Date),
      label = year(Date),
      linewidth = as.numeric(year(Date) == 2022)) +
  geom_line(data = discharge_df_cities2) +
  scale_color_discrete(type = c("grey50", "red")) +
  scale_linewidth() +
  geom_text_repel(
    data = discharge_df_cities %>% filter(yday(Date) == yday(max(Date))),
    color = "gray50",
    size = 3,
    direction = "y",
    nudge_x = 20,
    hjust = 0
  ) +
  scale_linewidth(range = c(0.5, 1)) +
  facet_wrap(vars(city), scales = "free_y") +
  labs(y = "Cumulative discharge anomaly [m3s-1]",
       x = "Date",
       color = "Year",
       fill = "Year") +
  theme(legend.justification = c(1, 0),
        legend.position = "none")
```

## Spatial Discharge ##

```{r}

dddf <- discharge_df %>%
  ## filter(Date >= as.Date("2022-07-01") & Date <= as.Date("2022-08-18")) %>%
  ## mutate(Date = cut(Date, "2 days")) %>%
  mutate(date_cut = cut(Date, datecuts, labels = datecutlabels),
         st_coordinates(discharge_df$geometry) %>% as.data.frame)

ocdf <- cbind(oder_cities, date_cut = factor(levels(cddf$date_cut)[3],
                                             levels = levels(cddf$date_cut)))

dddf %>%
  filter(!is.na(date_cut)) %>%
  ggplot() +
  geom_sf(data = river_net) +
  stat_summary_hex(aes(x = X, y = Y, z = discharge_anomaly),
                   fun = mean, na.rm = TRUE) +
  ## geom_point(aes(x = X, y = Y, color = Chlorophyll_mean)) +
  geom_sf(data = oder_cities) +
  geom_segment(data = ocdf,
               aes(xend = Inf, yend = after_stat(y), geometry = geometry),
               color = "gray70", stat = "sf_coordinates") +
  geom_label(data = ocdf,
             aes(label = name, geometry = geometry),
             x = Inf, hjust = 1, label.size = NA,
             color = "gray70", stat = "sf_coordinates") +
  scale_fill_gradient2(expression(atop("Discharge anomaly",
                                       paste("[", m^3, "/s]"))),
                       high = "#313695",
                       mid = "#ffffbf",
                       low = "#a50026") +
  ## scale_fill_viridis(expression("Chl. [" * mu * "g/l]"), direction = -1) +
  facet_wrap(vars(date_cut)) +
  ## facet_wrap(vars(Date)) +
  theme_void()

```

# Chlorophyll vs. discharge #
## Join chlorophyll data and discharge data ##

```{r}

chlorophyll_data_sf$location_idx <- discharge_location_idxs_lin

chlorophyll_discharge_df <- full_join(chlorophyll_data_sf,
          discharge_df %>%
            select(Date,
                   discharge_anomaly, discharge, discharge_perc,
                   location_idx) %>%
            filter(Date >= min(chlorophyll_data_sf$Date) &
                     Date <= max(chlorophyll_data_sf$Date)))
```


## Plot correlations ##

correlations on non-log data


## Plot correlations 2 ##

```{r}
cddf <- chlorophyll_discharge_df %>%
  mutate(date_cut = cut(Date, datecuts, labels = datecutlabels)) 

cities_annotation <- cbind(oder_cities, date_cut = cddf$date_cut[1])

## Discharge anomaly vs. cholorphyll
ggplot(cddf) +
  geom_vline(xintercept = oder_cities$distance_from_mouth / 1000, color = "gray70") +
  geom_text(data = cities_annotation,
            aes(x = distance_from_mouth / 1000, y = Inf, label = name),
            vjust = 1.1, hjust = 0, angle = -90, color = "gray70") +
  aes(x = distance_from_mouth / 1000) +
  geom_point(aes(y = Chlorophyll_mean, color = "Chlorophyll")) +
  geom_point(aes(y = discharge_anomaly, color = "Discharge anomaly")) +
  geom_smooth(aes(y = Chlorophyll_mean), color = "#b2df8a", span = 0.2, fill = NA) +
  geom_smooth(aes(y = discharge_anomaly), color = "#a6cee3", span = 0.2, fille = NA) +
  scale_x_reverse() +
  scale_y_continuous(
    limits = c(-400, 400),
    sec.axis = sec_axis(~ ., name = expression("Discharge anomaly [" * m ^ 3 * "/s]"))
  ) +
  scale_color_discrete("", type = c("#33a02c", "#1f78b4")) +
  labs(y = expression("Chlorophyll concentration [" * mu * "g/l]"),
       x = "Distance to river mouth [km]") +
  facet_wrap(vars(date_cut), nrow = 4) +
  theme_minimal() +
  theme(legend.position = c(0.7, 0),
        legend.justification = c(0.7, 0),
        legend.title = element_blank(),
        legend.margin = margin(0, 0, 0, 0),
        legend.box.margin = margin(0, 0, 0, 0),
        legend.background = element_rect(fill = "white", color = NA))

## Discharge anomaly percentile vs clorophyll concentration
ggplot(cddf) +
  geom_vline(xintercept = oder_cities$distance_from_mouth / 1000, color = "gray70") +
  geom_text(data = cities_annotation,
            aes(x = distance_from_mouth / 1000, y = Inf, label = name),
            vjust = 1.1, hjust = 0, angle = -90, color = "gray70") +
  aes(x = distance_from_mouth / 1000) +
  geom_point(aes(y = Chlorophyll_mean, color = "Chlorophyll")) +
  geom_point(aes(y = discharge_perc * 3, color = "Discharge percentile")) +
  geom_smooth(aes(y = Chlorophyll_mean), color = "#b2df8a", span = 0.2, fill = NA) +
  geom_smooth(aes(y = discharge_perc * 3), color = "#a6cee3", span = 0.2, fill = NA) +
  scale_x_reverse() +
  scale_y_continuous(
    limits = c(0, 300),
    sec.axis = sec_axis(~ . / 3, name = "Discharge anomaly percentile")
  ) +
  scale_color_discrete("", type = c("#33a02c", "#1f78b4")) +
  labs(y = expression("Chlorophyll concentration [" * mu * "g/l]"),
       x = "Distance to river mouth [km]") +
  facet_wrap(vars(date_cut), nrow = 4) +
  theme_minimal() +
  theme(legend.position = c(1, 1),
        legend.justification = c(1, 1),
        legend.title = element_blank(),
        legend.margin = margin(0, 0, 0, 0),
        legend.box.margin = margin(0, 0, 0, 0),
        legend.background = element_rect(fill = "white", color = NA))

```


```{r}
sessionInfo()
```
